---
title: "Laseri fat trajectories"
author: "Jouni Tuomisto"
date: "2021-03-24"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview and tasks

This project aims to create individual trajectories of fat mass change in individuals in the Laseri study.

### Tasks already performed

1. Create a Bayesian model that predicts the mean trajectory for girls and boys. 
2. First, use simple polynomial fit with three parameters ax^2 + bx + c.
3. Use multivariable normal for parameters a, b, c.
4. Use PFDWB for 1980-1992 and rasvapros01_lasklog for 2001.
5. R package sme checked but no longer available at CRAN, so other solutions are needed.

### To be done:

1. Decide on the function for fat percentage change.
2. Decide on the function for individual fat trajectory.
3. Consider using splines for modelling individual changes in fat percentage. Noora's master's thesis about smoothing splines for non-linear responses may be of help. Possibly useful articles https://doi.org/10.1214/ss/1038425655 https://doi.org/10.1198/1061860043010 
4. 

```{r preprocess}
library(reshape2)
library(tidyverse)

df <- read.csv("data/priv/rasvapopperus.csv")[-(11:85)] # Remove POP variables
# Each row is an individual
dfl <- melt(df, id.vars = c("tutkno80","SP","IKA80"), variable.name = "Parameter", value.name = "Result")
dfl$year[grepl("80",dfl$Parameter)] <- 1980
dfl$year[grepl("83",dfl$Parameter)] <- 1983
dfl$year[grepl("86",dfl$Parameter)] <- 1986
dfl$year[grepl("92",dfl$Parameter)] <- 1992
dfl$year[grepl("01",dfl$Parameter)] <- 2001
dfl$Parameter <- gsub("(80|83|86|92|01)","",dfl$Parameter)  
dfl$age <- dfl$year - 1980 + dfl$IKA80

ggplot(dfl[dfl$Parameter %in% c("PFDWB","rasvapros_lasklog"),], aes(x=age, y=Result, colour=IKA80, group=tutkno80))+geom_line()+facet_wrap(~SP)

tmp <- dfl[dfl$Parameter %in% c("PFDWB","rasvapros_lasklog"),]
tmp <- aggregate(tmp["Result"], by = tmp[c("age","SP","IKA80")], FUN=function(x) mean(x, na.rm = TRUE))
ggplot(tmp, aes(x=age, y=Result, colour=IKA80, group=IKA80))+geom_line()+facet_wrap(~SP)


```

## Hierarchical Bayes


```{r bayes}

  library(rjags) # JAGS
  library(MASS) # mvrnorm
  library(car) # scatterplotMatrix
  library(mcmcr) # as.mcarray
  
  N <- 100 # params$N # defined as render parameter
  
  # Hierarchical Bayes model.
  
  mod <- textConnection(
    "
    var a[2,3], fat[I,5], s[I], age[I]; # , b[2], c[2]
    model{
      taum ~ dgamma(1.0,1.0) # tau for mu
      tauf[1] ~ dgamma(1.0,1.0) # tau for fat percentage in girls
      tauf[2] ~ dgamma(1.0,1.0) # tau for fat percentage in boys
      a[1,1:3] ~ dmnorm(c(0,0,0), Omega) # girls
      a[2,1:3] ~ dmnorm(c(0,0,0), Omega) # boys

      for(i in 1:I) { 
        for(y in 1:length(YEAR)) {
          mu[i,y] ~ dnorm(a[s[i],1]*(YEAR[y]+age[i])^2 + a[s[i],2]*(YEAR[y]+age[i]) + a[s[i],3], taum)
          fat[i,y] ~ dnorm(mu[i,y] , tauf[s[i]])
          pred[i,y] ~ dnorm(mu[i,y] , tauf[s[i]])
        }
      }
    }
  ")

  jags <- jags.model(
    mod,
    data = list(
      YEAR = c(0,3,6,12,21), # Year since 1980
      age = df$IKA80, # Age at 1980
      fat = df[c("PFDWB80","PFDWB83","PFDWB86","PFDWB92","rasvapros01_lasklog")], # Fat percentage
      s = df$SP,
      I = nrow(df),
      Omega = diag(3)/1000
    ),
    n.chains = 4,
    n.adapt = 1000
  )
  
  update(jags, 1000)
  
  samps.c <- coda.samples(
    jags, 
    variable.names=c("a","tau"),
    n.iter=N*10,
    thin=10
  )
  
  samps.j <- jags.samples(
    jags, 
    variable.names=c("a","taum", "tauf", "mu","pred"),
    n.iter=N*10,
    thin=10
  )

  scatterplotMatrix(t(samps.j$a[1,,,1]), main="Polynomial parameters for girls")
  scatterplotMatrix(t(samps.j$a[2,,,1]), main="Polynomial parameters for boys")
  scatterplotMatrix(t(samps.j$tauf[,,1]), main="Precision for girls and boys")
    
#  plot(t(samps.c)) # This works for desktop but not when knitting. Why?

  age <- 3:30
  out <- data.frame()
  for(i in 1:100) {
    for(j in 1:2) {
    out <- rbind(out, data.frame(
      Iter = i,
      sex = j,
      x = age,
      y = samps.j$a[j,1,i,1]*age^2 + samps.j$a[j,2,i,1]*age + samps.j$a[j,3,i,1]
    ))
    }
  }

ggplot(out, aes(x=x,y=y,colour=Iter,group=Iter))+geom_line()+facet_wrap(~sex)

tmp <- apply(samps.j$pred[,,1,1], MARGIN=1:2, FUN=mean)
colnames(tmp) <- c("pred80","pred83","pred86","pred92","pred01")
tmp <- data.frame(df,tmp)

ggplot(tmp, aes(x=PFDWB80, y=pred80, colour=SP))+geom_point()

ggplot(tmp, aes(x=PFDWB86, y=pred86, colour=SP))+geom_point()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
