---
title: "LASERI-trajectories"
author: "Jouni Tuomisto"
date: '2021-03-31'
output:
  word_document: default
  html_document: default
params:
  N: 500
  c: 1900
  adjusted_theme_thl: yes
  run_bayes: yes
  run_preprocess: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This analysis produces a model for trajectories of human body burdens of persistent organic pollutants (POPs).

### Muistiinpanoja kokouksesta 2020-10-07

Rasvamassa on tärkeämpi kuin BMI, koska niitä on useammalta vuodelta lapsen kasvaessa (80, 83, 86). Lisäksi nimenomaan rasva vaikuttaa POPien kinetiikkaan (ja POPit rasva-aineenvaihduntaan?).

Rasvan määrää on arvioitu usealla eri kaavalla rasvapoimuista. Panu ja Noora tarkastavat parhaan laskentatavan. Myös vyötärönympäryksen ja leptiinipitoisuuden saaminen kaavaan tutkitaan.

Ritter on hyvä teoreettinen malli POP-kinetiikasta. Lisäksi ruotsalaisilla on jokin deskriptiivinen malli (mikä?).

Panu ja Noora kaivavat myös alkuperäiset pituus- ja painomittaukset 80, 83, 86. Näiden perusteella voisi suunnitella oman mallin, jolla kuvataan yksilöiden pituuden, painon ja rasvamäärän muutosta ajassa. Tähän voisi hyödyntää myös kasvukäyrien arvoja priorina. Pitääkö kehittää oma bayes-malli?

Hypoteesia pitäisi täsmentää. Tällä hetkellä ilmassa on useita, mutta mikä on se olennainen?

* Onko POP-pitoisuuksilla yhteyttä sairauksiin (kun vakioidaan rasvan suhteen)?
* Selittääkö POPien yleistyminen obesiteettia? (Tätä on kyllä todella vaikea testata, koska meillä ei ole dataa energiansaannista ja liikunnan määrästä)
* Mitkä ovat POP-pitoisuuden determinantit? (Tämä ei ole ensisijainen kysymys.)
* Miten huomioidaan ruumiin koko, joka on olennainen determinantti sairauksien synnyssä?
* Kemikaalikoktailien tutkiminen on olennainen osa alkuperäistä tutkimussuunnitelmaa. (Miten tarkalleen ottaen?)
* Onko vierasainemetabolia sekoittava tekijä? Joistakin POPeista metabolia tunnetaan aika hyvin. Lisäksi on SNIP-määrityksiä, joista voidaan tehdä genotyyppausta CYP-entsyymien suhteen. Pitäisi kehittää jokin hypoteesi jota tutkitaan. Panulla on tästä jo jokin raportti.

## Data preprocessing and preliminary analyses

```{r data}
library(foreign)
library(psych)

df <- read.spss("~/R/dioxdisthuman/data/priv/VS _Henkilökohtaiset_POP-trajektorit_LASERI-aineistossa/updated_korjattu_LASERI_data_2020_05_WORKCOPY.sav", to.data.frame = TRUE)

#colnames(df)[grep("PI",toupper(colnames(df)))]

units <- read.csv("POP_units.csv")

# Get normalised (per fat?) concentrations
tmp <- df[grepl("_norm",colnames(df))][-2] # remove alfa_HCH01, which contains no data
colnames(tmp) <- gsub("_norm","",colnames(tmp))
# Compare 1980 (80) and 2001 (01) in pairwise correlations
tmp <- tmp[sort(colnames(tmp))]
pairs.panels(tmp[1:8])
pairs.panels(tmp[8:15])
pairs.panels(tmp[15:22])
pairs.panels(tmp[22:29])
pairs.panels(tmp[29:36])

#colnames(tmp)
# [1] "HCB01"              "beta_HCH01"         "Trans_nonakloori01"
# [4] "pp_DDT01"           "pp_DDE01"           "PCB7401"           
# [7] "PCB9901"            "PCB11801"           "PCB15301"          
#[10] "PCB13801"           "PCB15601"           "PCB18701"          
#[13] "PCB18301"           "PCB18001"           "PCB17001"          
#[16] "BDE4701"            "kotiniini01"        "PeCB80"            
#[19] "HCB80"              "alfa_HCH80"         "beta_HCH80"        
#[22] "gamma_HCH80"        "Trans_nonakloori80" "pp_DDT80"          
#[25] "pp_DDE80"           "PCB7480"            "PCB9980"           
#[28] "PCB11880"           "PCB15380"           "PCB13880"          
#[31] "PCB15680"           "PCB18780"           "PCB18380"          
#[34] "PCB18080"           "PCB17080"           "kotiniini80"       

POP <- c(
  "HCB", # https://en.wikipedia.org/wiki/Hexachlorobenzene
  "alfa_HCH", # https://en.wikipedia.org/wiki/Alpha-Hexachlorocyclohexane
  "beta_HCH",
  "gamma_HCH", # also known as lindane
  "Trans_nonakloori", # https://en.wikipedia.org/wiki/Chlordane
  "pp_DDT", # para,para'-DDT https://en.wikipedia.org/wiki/DDT
  "pp_DDE", 
  "PCB74", # ug/kg fat, https://en.wikipedia.org/wiki/Polychlorinated_biphenyl
  "PCB99",
  "PCB118",
  "PCB153",
  "PCB138",
  "PCB156",
  "PCB187",
  "PCB183",
  "PCB180",
  "PCB170",
  "BDE47", # https://en.wikipedia.org/wiki/Pentabromodiphenyl_ether
  "kotiniini" # https://en.wikipedia.org/wiki/Cotinine
)

# Take all-male panel of interesting variables
df <- cbind(df[c(
    "SYNTVUOSI", # birth year
    "BMI80", # body mass index in 1980
    "BMI83", # body mass index in 1983
    "BMI86", # body mass index in 1986
    "BMI01", # body mass index in 2001
    "Paino01", # body weight kg in 2001
    "bodyfat80", # body fat % in 1980
    "bodyfat83", # body fat % in 1983
    "bodyfat86", # body fat % in 1986
    "rasva80", # serum fat content mg/ml in 1980
    "rasva01", # serum fat content mg/ml in 2001
    "lihatv80", # consumption of meat and meat products g/d in 1980
    "KALAV80", # consumption of fish and fish products g/d in 1980
    "FISHTOT_FFQ07", # total fish consumption, g/d in 2007 food frequency questionnaire
    "pai01k", # physical activity index in 2001
    "ADIPO01", # adiponectin hormone in serum, ug/ml in 2001
    "leptin01" # leptin hormone concentration ng/ml in 2001
  )],
  tmp
)[df$SP=="Poika", ]
# Very few observations in bodyfat92, Paino80, leptin80 --> omit
# It would be good to obtain these variables:
# ENERC07 energy consumption kJ/d from 2007 FFQ
# FAT07 fat consumption g/d from 2007 FFQ

pairs.panels(df[1:7])
pairs.panels(df[7:13])
pairs.panels(df[13:19])

### Select the population for analysis

dfs <- na.omit(df[c(1:2,5:6,14,31:32)])

## Indices used

AGE <- 0:40
H <- length(AGE)
YEAR <- 1960:2020
I <- nrow(dfs)
SYNT <- dfs$SYNTVUOSI

## Data used

U <- 365 # unit conversion factor: 1/d --> 1/a
A <- 0.9 # fraction, absorption fraction
N <- params$N
b <- 0.07 # 1/a, rate of decrease of the POP in the environment and intake
c <- params$c # year of peak exposure
P <- rep(1,H) # age-specific intake relative to lifetime average
# Estimated from the Bayesian model:
## k <- 0.35 # 1/a, elimination constant of the POP
## a <- 1.1 # ug/kg/d, average intake at the peak exposure

conc <- dfs[1:I,c("PCB15380","PCB15301")]
fish <- dfs$FISHTOT_FFQ07[1:I]

# https://www.kasvuseula.fi/
W <- t(matrix(rep(c(5,10,13,15,17,20,22,26,28,31,35,38,44,48,55,60,65,68,70,71,72,
                    rep(75,10),rep(80,10)),I),
              nrow=H, dimnames=list(Age=AGE, Individual = 1:I)))

# A simplistic model uses the population average as default until the weight reaches the
# observed value of that individual. Then, actual value is used to replace 80 kg (the max default),
# if larger. This leads to situation where people suddenly gain weight when at 31 or at 2001.
W <- pmin(W, dfs$Paino01, na.rm = TRUE)
W <- ifelse(W==80, pmax(W, dfs$Paino01, na.rm=TRUE), W)
tmp <- dfs$SYNTVUOSI + matrix(rep(AGE,each=I),nrow=I)>=2001
W <- ifelse(!is.na(dfs$Paino01) & tmp, dfs$Paino01, W)

# A simplistic model that assumes that average BMI is equal to lifetime fat percentage.
M <- W * ifelse(is.na(dfs$BMI01), dfs$BMI80, (dfs$BMI80 + dfs$BMI01)/2) / 100

#aggregate(df$Paino01, by=df[c("Age80")], FUN=function(x) mean(x, na.rm=TRUE))

```
The following conclusions can be made from the graphs above

* 1. PCBs correlate very well with each other.
* 2. PCB correlation is much stronger between congeners at the same year  (R>0.8) than the same congener 21 years apart (R = 0.2 - 0.4).
* 3. DDT and DDE congeners correlate moderately with each other and across years (R ca. 0.4)
* 4. Trans-nonachlor correlates well with DDE of the same year (R 0.45-0.6) and moderately across years (R ).
* 5. HCH congeners correlate only with alpha-HCH in 1980 (R 0.3 - 0.5) but not across years nor beta vs gamma.
* 6. HCB correlates well with alpha-HCH in 1980 but poorly across other congeners or years.
* 7. Cotinine correlates poorly across years (R = 0.2) and not at all with POPs.

## Deterministic data analyses

```{r deterministic}
library(ggplot2)
mimax <- c(min(df[c("PCB15380","PCB15301")]),max(df[c("PCB15380","PCB15301")]))
ggplot()+
  geom_point(data=df, aes(x=PCB15380, y=PCB15301, size=Paino01, colour=SYNTVUOSI))+
  geom_line(data=data.frame(x=mimax, y=mimax), aes(x=x, y=y),color="red")+
  scale_x_log10()+scale_y_log10()+
  labs(title="Correlation of individuals' PCB measurements 1980 and 2001")

df$id <- 1:nrow(df)
df$Paino80 <- NA
df$bodyfat01 <- NA
long <- reshape(df[c("SYNTVUOSI","id","PCB15380","PCB15301","BMI80","BMI01","Paino80","Paino01","bodyfat80","bodyfat01")],
                idvar=c("SYNTVUOSI","id"),
                varying=list(
                  PCB=c("PCB15380","PCB15301"),
                  BMI= c("BMI80","BMI01"),
                  weight=c("Paino80","Paino01"),
                  fat=c("bodyfat80","bodyfat01")
                ),
                v.names =c("PCB","BMI","weight","fat"),
                timevar="year",
                times=c(1980,2001),
                direction="long")

ggplot(long, aes(x=year, y=PCB, colour=SYNTVUOSI, group=id))+geom_line()+
  scale_y_log10()

ggplot(aggregate(long["PCB"], by=long[c("SYNTVUOSI","year")], FUN=function(x) mean(x, na.rm=TRUE)),
       aes(x=year, y=PCB, colour=SYNTVUOSI, group=SYNTVUOSI))+geom_line()

long$age <- long$year - long$SYNTVUOSI
ggplot(aggregate(long["PCB"], by=long[c("SYNTVUOSI","age")], FUN=function(x) mean(x, na.rm=TRUE)),
       aes(x=age, y=PCB, colour=factor(SYNTVUOSI), group=SYNTVUOSI))+geom_line()

cat("Concentrations in 20-year-old people decrease", signif((1-(40/120)^(1/(1977-1962)))*100,2), " % per year\n")

ggplot(aggregate(long["BMI"], by=long[c("SYNTVUOSI","age")], FUN=function(x) mean(x, na.rm=TRUE)),
       aes(x=age, y=BMI, colour=factor(SYNTVUOSI), group=SYNTVUOSI))+geom_line()

```
## Linear regressions to analyse impacts of variables

```{r regression}



# Remove the least important variables from the model
#reg <- lm(PCB15380 ~ SYNTVUOSI + BMI80 + bodyfat80 + KALAV80 + lihatv80, data=df)
#reg <- lm(PCB15380 ~ SYNTVUOSI + BMI80 + bodyfat80 + KALAV80, data=df)
reg <- lm(PCB15380 ~ SYNTVUOSI + BMI80 + bodyfat80, data=df)
summary(reg)

#reg <- lm(PCB15301 ~ SYNTVUOSI + BMI80 + BMI01 + Paino01 + PCB15380 + bodyfat80 + KALAV80 + lihatv80 + FISHTOT_FFQ07, data=df) 
#reg <- lm(PCB15301 ~ SYNTVUOSI + BMI80 + BMI01 + Paino01 + PCB15380 + bodyfat80 + KALAV80 + FISHTOT_FFQ07, data=df) 
#reg <- lm(PCB15301 ~ SYNTVUOSI + BMI80 + BMI01 + Paino01 + PCB15380 + bodyfat80 + FISHTOT_FFQ07, data=df) 
#reg <- lm(PCB15301 ~ SYNTVUOSI + BMI80 + BMI01 + Paino01 + PCB15380 + FISHTOT_FFQ07, data=df) 
reg <- lm(PCB15301 ~ SYNTVUOSI + BMI80 + BMI01 + PCB15380 + FISHTOT_FFQ07, data=df) 
reg <- step(reg, direction = "both")
summary(reg)

det1 <- "SYNTVUOSI + BMI80 + BMI01 + Paino01 + bodyfat80"
det2 <- " + KALAV80 + lihatv80 + FISHTOT_FFQ07"
df$alfa_HCH01 <- NA
df$BDE4780 <- NA
df$gamma_HCH01 <- NA

out <- data.frame()
for(j in POP) {
  for(i in c("80","01")) {
    for(k in 1:2) {
      if(!all(is.na(df[[paste0(j,i)]]))) {
      reg <- lm(as.formula(paste0(j, i," ~ ", c(det1, paste0(det1, det2))[k])), data = df)
      out <- rbind(out,
                   cbind(
                      congener=j,
                      year=i,
                      formula = c("short","long")[k],
                      as.data.frame(summary(reg)[[4]]),
                      N = summary(reg)[[7]][2]+1,
                      R2 = summary(reg)[[8]][1]
                   )
      )
  }}}
}

tmp <- aggregate(out["R2"], by=out[c("congener","year","formula")],FUN=mean)
tmp <- tmp[order(-tmp$R2),]
ggplot(tmp, aes(x=congener, y=R2, colour=formula, shape=year))+geom_point(size=2)+
  coord_flip()+
  labs(title="Regression fit (R^2) of a pollutant with more or less\n explanatory variables at different years")

```

For PCB15380, we can see that

1. KALAV80 and lihatv80 bring little if any predictive power to the regression model
2. FISHTOT_FFQ07 seems to have some predictive power


## Bayesian hierachical model for POP kinetics

```{r bayes}
# This was forked from Dioxdistboys_individual_rows.Rmd and
# originally from code Op_en3104/bayes on page [[EU-kalat]]

## This code builds an iterative model with varying intake
## https://github.com/jtuomist/dioxdisthuman/wiki#iterative-model-with-varying-intake-primary-choice

library(OpasnetUtils)
library(reshape2)
library(rjags) # JAGS
library(MASS) # mvrnorm
library(car) # scatterplotMatrix
library(mcmcr) # as.mcarray

N <- params$N # defined as render parameter
  
###########################3# Hierarchical Bayes model.
## The model estimates POP body burden (B) trajectories at individual level. It calculates iteratively.
## It calculates B yearly and calculates concentrations C = B / M where M is body fat mass for those years
## that have C observations available.

#  The variables in the model:
# See https://github.com/jtuomist/dioxdisthuman/wiki/Modelling-POP-concentrations-in-humans#answer-summary
  
#    B = body burden of the POPs at individual level with indices i = individual, t = age, y = calendar year
#    k = elimination constant elimination constant for each POP from the body. This can also be evaluated in the model, using Ritter et al estimates as priors. Or are there good published k estimates available?
#    U = 365, constant for unit conversion (1/day --> 1/year)
#    A = 0.9, absorption fraction of POPs from the gut. Typically close to 1, but more precise POP-specific data may be available.
#    W = body weight of an individual. Comes from LASERI data.
#    E = reference daily exposure at year y using parameters a and b.
#    a = exposure at year 2000
#    b = 0.07 1/a, rate of decrease of the POP in the environment and intake
#    P = age-dependent factor to adjust for breast feeding and other age-dependent POP exposure. Should we include individual variation of amount of breast feeding? Yes we should, if that makes a difference for conclusions.
#    C = observed concentrations of the POPs at individual level (i) at different ages (t).
#    M = body fat amount (in kg) of an individual. Comes from a separate model.

  mod <- textConnection(
    "
    var k, a, b, U, E[61], C[I,H], B[I,H], M[I,H], C1, C2, conc[I,2], SYNT[I], Efish[I],
    B1[I], B2[I], B3[I], B4[I], B5[I], B6[I], B7[I], B8[I], B9[I], B10[I], B11[I],
    B12[I], B13[I], B14[I], B15[I], B16[I], B17[I], B18[I], B19[I], B20[I], B21[I], B22[I],
    B23[I], B24[I], B25[I], B26[I], B27[I], B28[I], B29[I], B30[I], B31[I], B32[I], B33[I],
    B34[I], B35[I], B36[I], B37[I], B38[I], B39[I], B40[I], B41[I];
    model{
      k ~ dnorm(0.3, 0.1)
      a ~ dnorm(1, 0.1)
#      b ~ dnorm(0.1, 0.001) # assumed constant at the moment
      tau[1] ~ dgamma(1.0,1.0) 
      tau[2] ~ dgamma(1.0,1.0) 
#      bfish ~ dnorm(1.0,0.001) 
      bfish <- 0
      Efish[] <- fish[] * bfish
      for(year in YEAR) {
        E[year-1959] <- a*exp(-b*(year-2000))
      }
      B2[] <-  B1[] * (1 - k) + U * A * W[,1] * (E[SYNT[]-1960+1] + Efish[]) * P[1]
      B3[] <-  B2[] * (1 - k) + U * A * W[,2] * (E[SYNT[]-1960+2] + Efish[]) * P[2]
      B4[] <-  B3[] * (1 - k) + U * A * W[,3] * (E[SYNT[]-1960+3] + Efish[]) * P[3]
      B5[] <-  B4[] * (1 - k) + U * A * W[,4] * (E[SYNT[]-1960+4] + Efish[]) * P[4]
      B6[] <-  B5[] * (1 - k) + U * A * W[,5] * (E[SYNT[]-1960+5] + Efish[]) * P[5]
      B7[] <-  B6[] * (1 - k) + U * A * W[,6] * (E[SYNT[]-1960+6] + Efish[]) * P[6]
      B8[] <-  B7[] * (1 - k) + U * A * W[,7] * (E[SYNT[]-1960+7] + Efish[]) * P[7]
      B9[] <-  B8[] * (1 - k) + U * A * W[,8] * (E[SYNT[]-1960+8] + Efish[]) * P[8]
      B10[] <-  B9[] * (1 - k) + U * A * W[,9] * (E[SYNT[]-1960+9] + Efish[]) * P[9]
      B11[] <-  B10[] * (1 - k) + U * A * W[,10] * (E[SYNT[]-1960+10] + Efish[]) * P[10]
      B12[] <-  B11[] * (1 - k) + U * A * W[,11] * (E[SYNT[]-1960+11] + Efish[]) * P[11]
      B13[] <-  B12[] * (1 - k) + U * A * W[,12] * (E[SYNT[]-1960+12] + Efish[]) * P[12]
      B14[] <-  B13[] * (1 - k) + U * A * W[,13] * (E[SYNT[]-1960+13] + Efish[]) * P[13]
      B15[] <-  B14[] * (1 - k) + U * A * W[,14] * (E[SYNT[]-1960+14] + Efish[]) * P[14]
      B16[] <-  B15[] * (1 - k) + U * A * W[,15] * (E[SYNT[]-1960+15] + Efish[]) * P[15]
      B17[] <-  B16[] * (1 - k) + U * A * W[,16] * (E[SYNT[]-1960+16] + Efish[]) * P[16]
      B18[] <-  B17[] * (1 - k) + U * A * W[,17] * (E[SYNT[]-1960+17] + Efish[]) * P[17]
      B19[] <-  B18[] * (1 - k) + U * A * W[,18] * (E[SYNT[]-1960+18] + Efish[]) * P[18]
      B20[] <-  B19[] * (1 - k) + U * A * W[,19] * (E[SYNT[]-1960+19] + Efish[]) * P[19]
      B21[] <-  B20[] * (1 - k) + U * A * W[,20] * (E[SYNT[]-1960+20] + Efish[]) * P[20]
      B22[] <-  B21[] * (1 - k) + U * A * W[,21] * (E[SYNT[]-1960+21] + Efish[]) * P[21]
      B23[] <-  B22[] * (1 - k) + U * A * W[,22] * (E[SYNT[]-1960+22] + Efish[]) * P[22]
      B24[] <-  B23[] * (1 - k) + U * A * W[,23] * (E[SYNT[]-1960+23] + Efish[]) * P[23]
      B25[] <-  B24[] * (1 - k) + U * A * W[,24] * (E[SYNT[]-1960+24] + Efish[]) * P[24]
      B26[] <-  B25[] * (1 - k) + U * A * W[,25] * (E[SYNT[]-1960+25] + Efish[]) * P[25]
      B27[] <-  B26[] * (1 - k) + U * A * W[,26] * (E[SYNT[]-1960+26] + Efish[]) * P[26]
      B28[] <-  B27[] * (1 - k) + U * A * W[,27] * (E[SYNT[]-1960+27] + Efish[]) * P[27]
      B29[] <-  B28[] * (1 - k) + U * A * W[,28] * (E[SYNT[]-1960+28] + Efish[]) * P[28]
      B30[] <-  B29[] * (1 - k) + U * A * W[,29] * (E[SYNT[]-1960+29] + Efish[]) * P[29]
      B31[] <-  B30[] * (1 - k) + U * A * W[,30] * (E[SYNT[]-1960+30] + Efish[]) * P[30]
      B32[] <-  B31[] * (1 - k) + U * A * W[,31] * (E[SYNT[]-1960+31] + Efish[]) * P[31]
      B33[] <-  B32[] * (1 - k) + U * A * W[,32] * (E[SYNT[]-1960+32] + Efish[]) * P[32]
      B34[] <-  B33[] * (1 - k) + U * A * W[,33] * (E[SYNT[]-1960+33] + Efish[]) * P[33]
      B35[] <-  B34[] * (1 - k) + U * A * W[,34] * (E[SYNT[]-1960+34] + Efish[]) * P[34]
      B36[] <-  B35[] * (1 - k) + U * A * W[,35] * (E[SYNT[]-1960+35] + Efish[]) * P[35]
      B37[] <-  B36[] * (1 - k) + U * A * W[,36] * (E[SYNT[]-1960+36] + Efish[]) * P[36]
      B38[] <-  B37[] * (1 - k) + U * A * W[,37] * (E[SYNT[]-1960+37] + Efish[]) * P[37]
      B39[] <-  B38[] * (1 - k) + U * A * W[,38] * (E[SYNT[]-1960+38] + Efish[]) * P[38]
      B40[] <-  B39[] * (1 - k) + U * A * W[,39] * (E[SYNT[]-1960+39] + Efish[]) * P[39]
      B41[] <-  B40[] * (1 - k) + U * A * W[,40] * (E[SYNT[]-1960+40] + Efish[]) * P[40]
      
      C[,1] <- B1 / M[,1]
      C[,2] <- B2 / M[,2]
      C[,3] <- B3 / M[,3]
      C[,4] <- B4 / M[,4]
      C[,5] <- B5 / M[,5]
      C[,6] <- B6 / M[,6]
      C[,7] <- B7 / M[,7]
      C[,8] <- B8 / M[,8]
      C[,9] <- B9 / M[,9]
      C[,10] <- B10 / M[,10]
      C[,11] <- B11 / M[,11]
      C[,12] <- B12 / M[,12]
      C[,13] <- B13 / M[,13]
      C[,14] <- B14 / M[,14]
      C[,15] <- B15 / M[,15]
      C[,16] <- B16 / M[,16]
      C[,17] <- B17 / M[,17]
      C[,18] <- B18 / M[,18]
      C[,19] <- B19 / M[,19]
      C[,20] <- B20 / M[,20]
      C[,21] <- B21 / M[,21]
      C[,22] <- B22 / M[,22]
      C[,23] <- B23 / M[,23]
      C[,24] <- B24 / M[,24]
      C[,25] <- B25 / M[,25]
      C[,26] <- B26 / M[,26]
      C[,27] <- B27 / M[,27]
      C[,28] <- B28 / M[,28]
      C[,29] <- B29 / M[,29]
      C[,30] <- B30 / M[,30]
      C[,31] <- B31 / M[,31]
      C[,32] <- B32 / M[,32]
      C[,33] <- B33 / M[,33]
      C[,34] <- B34 / M[,34]
      C[,35] <- B35 / M[,35]
      C[,36] <- B36 / M[,36]
      C[,37] <- B37 / M[,37]
      C[,38] <- B38 / M[,38]
      C[,39] <- B39 / M[,39]
      C[,40] <- B40 / M[,40]
      C[,41] <- B41 / M[,41]
      
      for(i in 1:I) { 
        conc[i,1] ~ dnorm(C[i,1980 - SYNT[i]+1], tau[1]) # Age 0 is at position 1
        conc[i,2] ~ dnorm(C[i,2001 - SYNT[i]+1], tau[2])
        pred[i,1] ~ dnorm(C[i,1980 - SYNT[i]+1], tau[1])
        pred[i,2] ~ dnorm(C[i,2001 - SYNT[i]+1], tau[2])
      }
    }
  ")

  jags <- jags.model(
    mod,
    data = list(
      YEAR = YEAR,
      W = W[1:I,1:H],
      M = M[1:I,1:H],
      H = H,
      I = I,
      U = U,
      A = A,
      b = b,
#      c = c,
      P = P[1:H],
      SYNT = SYNT,
      conc = conc,
      fish = fish,
      B1 = rep(0,I)
    ),
    n.chains = 4,
    n.adapt = 300
  )
  
  samps.c <- coda.samples(
    jags, 
    variable.names=c("k","a","b","tau","bfish"), 
    n.iter=N*10,
    thin=10
  )
  
  samps.j <- jags.samples(
    jags, 
    variable.names=c("k","a","b","tau","E","pred","bfish"), 
    n.iter=N*10,
    thin=10
  )

  #scatterplotMatrix(samps.j$k, main="Elimination constant k")
    
  plot(samps.c)

#}

```
* k: elimination constant (1/a)
* a: maximum exposure at the year of peak production (ug/kg/d)
* b: nominal rate of environmental concentration decrease (1/a)
* tau: precision for individual concentration estimate at year x (1980 or 2001)
* E: exposure at a calendar year y (between 1960 and 2020)
* pred: predicted individual concentration (ug/kg fat)

## Results from the Bayesian kinetic model

```{r results}

k <- mean(samps.j$k)
a <- mean(samps.j$a)
b <- mean(samps.j$b)

E <- a*exp(-b*(YEAR-2000))
B <- matrix(0, nrow=I, ncol=H)
for(age in 2:H) {
    B[,age] <-  B[,age-1] * (1 - k) + U * A * W[,age-1] * E[SYNT-1963+age] * P[age]
}
C <- B / M

ggplot(data.frame(year=YEAR,exposure=E), aes(x=year, y=E))+geom_line()+
  labs(title="Exposure level by time",
       y="Exposure (ug/kg/d)")

dat <- melt(B)#[seq(1,550, 10),])
dat <- merge(dat, cbind(Var1= 1:I, SYNT=SYNT))

ggplot(dat, aes(x=Var2, y=value, colour=SYNT, group=Var1))+geom_line()+
  labs(title="Body burdens in individuals",
       y="Total body burden (ug)")

ggplot(aggregate(dat["value"], by=dat[c("Var2","SYNT")], FUN=function(x) mean(x, na.rm=TRUE)), 
       aes(x=Var2, y=value, colour=factor(SYNT), group=SYNT))+geom_line()+
  labs(title="Body burden average by age group",
       y="Total body burden (ug)")

dat <- melt(C)#[seq(1,550, 10),])
dat <- merge(dat, cbind(Var1= 1:I, SYNT=SYNT))

ggplot(dat, aes(x=Var2, y=value, colour=SYNT, group=Var1))+geom_line()+
  labs(title="Concentrations in individuals",
       y="Concentration (ug/kg fat)")

ggplot()+
  geom_line(data=aggregate(dat["value"], by=dat[c("Var2","SYNT")], FUN=function(x) mean(x, na.rm=TRUE)), 
       aes(x=Var2, y=value, colour=factor(SYNT), group=SYNT))+
  geom_point(data=aggregate(long["PCB"],by=long[c("SYNTVUOSI","age")],FUN=function(x) mean(x, na.rm=TRUE)),
             aes(x=age, y=PCB,colour=factor(SYNTVUOSI)))+geom_point()+
  labs(title="Concentration average by age group",
       y="Concentration (ug/kg fat")

tst <- rowMeans(samps.j$pred[,,,], dims=2)
tst <- data.frame(
  dfs[c("PCB15380")],
  PCB15380_pred = tst[,1],
  dfs[c("PCB15301")],
  PCB15301_pred = tst[,2]
)
pairs.panels(tst)

```
In conclusion, the model predicts OK the agegroup averages but does not predict individual variation. The correlation between measured and predicted PCB153 in 2001 are actually negative.

In addition: why do the predicted values form two subpopulations when 1980 and 2001 are scatterplotted?
